---
title: "Handling OUCRU publications"
format: html
editor: visual
editor_options: 
  chunk_output_type: console
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(cache = FALSE, message = FALSE, warning = FALSE)
```

Packages needed:

```{r}
library(readxl)
library(dplyr)
library(rcrossref)
```

Reading the year 2022 from OUCRU excel file as an example:

```{r}
pubs22 <- read_excel("2022 OUCRU Publications (Mar20).xlsx")
```

`pubs22` contains the data from the first sheet of the excel file (here showing
only the first 16 columns of the first 10 publications as an example):

```{r}
pubs22
```

From `pubs22` we want to extract the DOIs:

```{r}
dois <- pull(pubs22, "DOI")
```

which looks like this (just showing the first few DOIs as an example):

```{r}
head(dois)
```

Then you can use this list of DOIs to query the complete information of each
publication through [Crossref](https://www.crossref.org) online:

```{r eval = FALSE}
p22 <- cr_works(dois)
```

```{r include = FALSE}
p22 <- readRDS("p22.rds")
```

You can see the amazing quantity of information that is available for each
publication. Here a full list of the variables (no less than 36!):

```{r}
names(p22$data)
```

This is basically what it looks like (again, showing here only the first 8
fields of the first 10 publications):

```{r}
p22$data
```

Some of these variables are very interesting, such as `is.referenced.by.count`
that tells you how often the publication has been cited as of today! This is a
key metric. Also, interestingly, there is a variable named `subject` that gives
you the broad topic of the publication. There is also a `funder` variable that
tells you what funded the study. For example, for the 3rd publication in this
list we can see that it is Wellcome:

```{r}
p22$data[3, "funder"][[1]][[1]]$name
```

But probably more interesting to you is what relates to the authors information,
where given and family names are (1) full and (2) clearly distinguished, and the
affiliations are given too. Here an example for the first publication:

```{r}
p22$data[1, "author"][[1]]
```

With the full given and family names here I believe you can fairly easily
construct the hash table that will ensure consistency. Indeed, the most common
inconsistency that happens with Vietnamese authors is that they sometimes switch
given and family names. So, you can construct the hash table by trying both
options, which takes no time with a bit of computer code. One point to note is
that even if on the publication itself initials are usually used for the given
names, in the actually data bases such as the one that we queried here
([Crossref](https://www.crossref.org)), you'll always find the full names.
Hence, with the DOIs and a little bit of code you can actually retrieve much
more information than what is in the excel file that you sent me, in a tiny
fraction of the time that it took to build this excel file in the first place!
That's why I'm basically recommending OUCRU to forget about this excel file and
just make sure they have the DOI which is the only thing that really matters.
